---
title: "02 Random Forest"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries and what not

```{r message=FALSE}
## Base RF model
set.seed(42)

library(ggplot2)  # for autoplot() generic
library(dplyr)
library(sf)
library(ranger)
library(vip)
library(pdp)
library(iml)
```

## Data processing

```{r}
dat_sf <- st_read("../data/glacier_clim.shp")
dat <- dat_sf %>%
  st_drop_geometry()

dat <- dat %>% 
  dplyr::select(mb_mwea, area_km2, z_min, z_med, z_max, z_aspct, z_slope,
                tau,
                t2m_18, t2m_d, tp_18, tp_d)

dat2 <- dat[sample(1:nrow(dat), 500), ]
```

## Build model

```{r}
mb_rf <- ranger(mb_mwea ~ ., dat2, 
                importance = "permutation")
```

## Variable importance

```{r}
vip(mb_rf)
```

## Partial dependency

```{r}
pred_wrapper <- function(object, newdata) {
  # p <- predict(object, data = newdata)$predictions[, 1L, drop = TRUE]
  p <- predict(object, data = newdata)$predictions
  #c("avg" = mean(p), "avg-1sd" = mean(p) - sd(p), "avg+1sd" = mean(p) + sd(p))
  c("avg" = mean(p))
}

pd1 <- partial(mb_rf, "t2m_d", 
               train = dat2, 
               pred.fun = pred_wrapper, grid.resolution = 30)
pd2 <- partial(mb_rf, "tp_d", 
               train = dat2, 
               pred.fun = pred_wrapper, grid.resolution = 30)

pdp1 <- autoplot(pd1) + 
  theme_light() +
  labs(x = "dT2M", y = "Partial dependence") +
  theme(legend.position = "none")
pdp2 <- autoplot(pd2) + 
  theme_light() +
  labs(x = "dP", y = "Partial dependence") +
  theme(legend.position = "none")
grid.arrange(pdp1, pdp2, nrow = 1)  # display plots side by side

# pdp3 <- partial(mb_rf, pred.var = c("t2m_d", "t2m_18"), 
#                 chull = TRUE, grid.resolution = 30,
#                 pred.fun = pred_wrapper, train = dat2) %>%
#   autoplot()
# print(pdp3)  # display plot
# 
# pdp4 <- partial(mb_rf, pred.var = c("t2m_d", "z_med"), 
#                 chull = TRUE, grid.resolution = 30,
#                 pred.fun = pred_wrapper, train = dat2) %>%
#   autoplot()
# print(pdp4)  # display plot
```

## Interaction values

Convert the ranger output to an iml object:

```{r}
pred <- function(model, newdata)  {
  results <- as.data.frame(predict(model, newdata)$predictions)
  return(results)
}

mod <- Predictor$new(mb_rf, 
                     data = dat[-which(names(dat) == "mb_mwea")],
                     y = dat$mb_mwea,
                     predict.function = pred)
```

### Importance

```{r}
imp <- FeatureImp$new(mod, loss = "mse")
plot(imp)
```

### Effect sizes

Calculate individual feature effects (PDs)  and plot them using a not at all controversial plotting method:

```{r}
effect <- FeatureEffects$new(mod, 
                             grid.size = 30, 
                             method = "pdp")
plot(effect)
```

In the `results` part of this object are the individual PD values. We can then get their SD values for the effect size:

```{r}
eff_size <- as.data.frame(lapply(effect$results, function(x) sd(x$.value)))
knitr::kable(eff_size)
```

#### Effect size by level

```{r}
effect <- FeatureEffect$new(mod, c("t2m_d", "z_med"),
                            method = "pdp")
plot(effect)
```

### Interactions

#### Total interactions

Total effect of interactions on each feature:

```{r}
ia <- Interaction$new(mod)
ia$print()
```

```{r}
plot(ia)
```

#### Two-way interactions

These need to be calculated for each variable in turn, but give the proportion of the effect explained by interactions. Here's an example for mean precip

```{r}
ia_2way <- Interaction$new(mod, feature = "t2m_d")
ia_2way$print()
```

```{r}
ia_2way$plot()
```

